Chasing \rainbows{} with the Habitable Worlds Observatory

### 2023-03-10 src/tex/ms.tex
[showyourwork] overleaf sync

Surface liquid water is the definition of planetary habitability: water is required for life as we know it, and life will only be remotely detectable on an exoplanet if it is present at the surface. Hence the search for life on exoplanets typically focuses on habitable planets. To first order, the presence of surface liquid water can be predicted by its instellation---the flux it receives from its star. The habitable zone (HZ) describes the range of semi-major axes for which an Earth twin could maintain surface liquid water around a given main sequence star \cite{kasting93}.  The HZ is popular among planet hunters because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis. Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.

@@

In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and are themselves non-Lambertian scatterers. \cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves. Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization. \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
\citet{Zugger_2011} note that ocean glint is best observed in near infrared opacity windows because of the reduced Rayleigh scattering at long wavelengths. This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}. Most recently, \cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds. The bottom line is that phase-dependent reflected light photometry and polarization of a directly-imaged exoplanet could establish its habitability.

>>>>

Surface liquid water is the definition of planetary habitability: water is required for life as we know it, and life will only be detectable on an exoplanet if it is present at the surface. Hence the search for life on exoplanets typically focuses on habitable planets. To first order, the presence of surface liquid water can be predicted by the flux it receives from its star. The habitable zone (HZ) describes the range of semi-major axes for which an Earth twin could maintain surface liquid water around a given main sequence star \citep{kasting93}.  The HZ is convenient because a planet can be deemed habitable based on solely on its orbital separation, but it remains an untested hypothesis. Ideally, we will empirically establish the presence of liquid water on an exoplanet before trying to interpret its putative biosignatures.

@@

In practice, clouds complicate the detection of ocean glint because they obscure the underlying surface and are themselves non-Lambertian scatterers. \cite{Zugger_2010} simulated photometric and polarized phase variations for planets including the effects of partial cloud coverage and ocean waves. Although the presence of clouds and atmospheric Rayleigh scattering weaken the signal of an underlying ocean, they again find that oceans are detectable and distinguishable from other sources of polarization. \cite{2010ApJ...721L..67R} compared observations of the Earth's scattering phase curve to comprehensive radiative transfer simulations; they found that the photometric brightening of Earth at crescent phases is mostly due to forward scattering from clouds, with a relatively small contribution from ocean glint.
\citet{Zugger_2011} note that ocean glint is best observed in near infrared opacity windows because of the reduced Rayleigh scattering at long wavelengths. This has the added advantage of minimizing the latitude--albedo effect, which could otherwise be a false-positive for photometric brightening at crescent phases due to glint \citep{2012ApJ...752L...3C}. Most recently, \cite{2019A&A...626A.129T,2022A&A...664A.172T} simulated the photometric and polarimetric signatures of terrestrial planets with inhomogeneous clouds. The bottom line is that a direct imaging mission sensitive to the phase-dependent reflected light photometry and polarization of an  exoplanet could establish its habitability.
